Staged Abstract Interpreters

### 2019-09-11 paper/50_stageabs.tex
typo

Our staging schema works by unfolding the interpreter over the abstract syntax
tree of the input program. In practice, however, it would suffer from code
explosion when analyzing (specializing) large programs, which increases the
compile time. If we generate next-stage programs running on JVM, such large
generated program would also potentially cause runtime GC overhead. In this
section, we present optimizations that largely mitigate these issues. For all
that, implementing these optimizations do not need to change the generic
interpreter.

@@

We implemented a big-step, compositional abstract interpreter in monadic
style, where \textit{compositional} means that every recursive call of our abstract
interpreter is applied to the proper substructures of the current syntactic
parameters \cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
specialization can be done by unfolding, as well as guarantees the termination
of the specialization procedure. It is also possible to specialize a small-step
operational abstract semantics through abstract compilation
\cite{Boucher:1996:ACN:647473.727587} -- as
\citet{Johnson:2013:OAA:2500365.2500604} presented for
optimizing Abstract Abstract Machines. However, the generated abstract
bytecode still requires another small-step abstract machine to execute, which is
an additional engineering effort.

>>>>

Our staging schema works by unfolding the interpreter over the abstract syntax
tree of the input program. In practice, however, it would suffer from code
explosion when analyzing (specializing) large programs, which increases the
compile time. If we generate next-stage programs running on JVM, such large
generated programs would also potentially cause runtime GC overhead. In this
section, we present optimizations that largely mitigate these issues. For all
that, implementing these optimizations do not need to change the generic
interpreter.

@@

We have implemented a big-step, compositional abstract interpreter in monadic
style, where \textit{compositional} means that every recursive call of our abstract
interpreter is applied to the proper substructures of the current syntactic
parameters \cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
specialization can be done by unfolding, as well as guarantees the termination
of the specialization procedure. It is also possible to specialize a small-step
operational abstract semantics through abstract compilation
\cite{Boucher:1996:ACN:647473.727587} -- as
\citet{Johnson:2013:OAA:2500365.2500604} presented for
optimizing Abstract Abstract Machines. However, the generated abstract
bytecode still requires another small-step abstract machine to execute, which is
an additional engineering effort.
